---
title: Objects & data types
---

::: {.callout-tip}
#### Learning objectives

- 
:::


## Context

We’ve seen examples where we entered data directly into a function. Most of the time we have data from elsewhere, such as a spreadsheet. In the previous section we created single objects. We’ll build up from this and introduce vectors and tabular data. We'll also briefly mention other data types, such as matrices, arrays.

## Objects

### Creating objects

Just running lines of code can be helpful if you only need an answer, but in order to do useful and interesting things, we often need to save values so we can work with them.

To do this, we *assign* values to *objects*. An object acts as a container for that value.

To create an object, we need to give it a name followed by the
assignment operator `<-`, and the value we want to give it, for example:

::: {.panel-tabset group="language"}
## R

```{r}
temperature <- 23
```

We can read the code as: the value `23` is assigned to the object `temperature`. Note that when you run this line of code the object you just created appears on your environment tab (top-right panel).

When assigning a value to an object, R does not print anything on the console. You can print the value by typing the object name on the console.

::: {.callout-important}
## The assignment operator

In R we use `<-` as the assignment operator. It assigns values on the right to objects on
the left. So, after executing:

In RStudio, typing <kbd>Alt</kbd> + <kbd>-</kbd> (push <kbd>Alt</kbd> at the same time as the <kbd>-</kbd> key) will write ` <- ` in a single keystroke on a PC, while typing <kbd>Option</kbd> + <kbd>-</kbd> (push <kbd>Option</kbd> at the same time as the <kbd>-</kbd> key) does the same on a Mac.

:::

Objects can be given almost any name such as `x`, `current_temperature`, or
`subject_id`. You want the object names to be explicit and short. There are some exceptions / considerations (see below).

::: {.callout-warning}
## Restrictions on object names

Object names can contain letters, numbers, underscores and periods. They *cannot start with a number nor contain spaces*. Different people use different conventions for long variable names, two common ones being:

Underscore: my_long_named_object

Camel case: myLongNamedObject

What you use is up to you, but be consistent. R is **case-sensitive** so `temperature` is different from `Temperature.`

* Some names are forbidden, because they are the names of fundamental functions in R (e.g.,
`if`, `else`, `for`, see
[here](https://stat.ethz.ch/R-manual/R-devel/library/base/html/Reserved.html)
for a complete list).
* Avoid using function names (e.g., `c`, `T`, `mean`, `data`, `df`, `weights`), even if allowed. If in doubt, check the help to see if the name is already in use.
* Avoid full-stops (`.`) within an object name as in `my.data`. Full-stops often have meaning in programming languages, so it's best to avoid them.
* Use consistent styling. In R, popular style guides are:
  * [tidyverse's](http://style.tidyverse.org/).
  * [Google's](https://google.github.io/styleguide/Rguide.xml)

**Whatever style you use, be consistent!**
:::
:::

### Using objects

Now that we have the `temperature` in memory, we can use it to perform operations. For example, this might the temperature in Celsius and we might want to calculate it to Kelvin.

To do this, we need to add `273.15`:

::: {.panel-tabset group="language"}
## R

```{r}
temperature + 273.15
```
:::

We can change an object's value by assigning a new one:

::: {.panel-tabset group="language"}
## R

```{r}
temperature <- 36
temperature + 273.15
```
:::

Finally, assigning a value to one object does not change the values of other objects. For example, let’s store the outcome in Kelvin into a new object `temp_K`:

::: {.panel-tabset group="language"}
## R

```{r}
temp_K <- temperature + 273.15
```
:::

Changing the value of `temperature` does not change the value of `temp_K`.

::: {.panel-tabset group="language"}
## R

```{r}
temperature <- 14
temp_K
```
:::

### Vectors

* LO: create vectors
* LO: operations with vectors
* LO: subsetting (indices, conditionals with TRUE/FALSE)

### Dealing with missing data

* LO: why is missing data important?
* LO: good practices of dealing with missing data

## Summary

::: {.callout-tip}
#### Key points

- 
:::
